Locality and Computational Reliability of Linear Response Calculations for Molecular Systems

### 2018-11-15 Response_referees.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

List of major changes apported to the paper:
\begin{itemize}
 \item The introduction has been deeply revised. In particular we have cut the first two
 paragraph concerning the differences between accuracy and precision.
 \item Added a comment after eq. (7) on the non-hermiticity of the response density and corrected
 the equation (11).
 \item Added a dedicated subsection (actually the IIA at page 4)  to discuss the localization properties of the FS states for static perturbations. 
 \item Added a panel in figure 2 that contain the plot of the real part of $\alpha(\omega)$ in calculated in different computational domain. This new plot aims to evidence that the localization
 properties of th FS affect, on equal footing, both the real and the imaginary part of the linear response functional.
 \item The relation between localization properties and features of the computational basis needed to provide precise results has been discussed in a stronger way in all the text. Major modifications under this perspective have been done in section II before the analysis of fig. 2 and....
\end{itemize}

### 2018-11-15 AnalyticStructure_main.tex
update

% Useful packages
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage{mathtools,braket}
\usepackage{tikz}
\usepackage{xifthen}
%\pgfplotsset{compat=1.15}
%\usepackage[markup=underline]{changes}
\usepackage[final]{changes}
\definechangesauthor{dal}
\definechangesauthor{gen}

@@

\section{Introduction}
\deleted[id=gen]{%
Since the days of their foundations, disciplines like Computational Physics or Quantum Chemistry always had to deal with the problem of the
computational reliability of the results. More specifically, a computational result may be considered reliable for various reasons.
For example, it may be in agreement with some previous observation or measurement that the employed computational model is susceptible to reproduce.
The model itself may in this way be classified as ``accurate''; it can then be employed to extract, \textit{in silico},
other quantities that might be then used to \emph{predict} or \emph{interpret} experiments.
Another concept of reliability is the concept of  ``precision'', which is more focused to the actual results
that a given computational model provides for a specific set of quantities, regardless of their potential agreement with
experimental data. The aim of precise computational approach is to reduce the \emph{computational} uncertainties
of quantities extracted out of a well-defined model, and provide reference results, to which other computer codes
employing the same model can be compared.
%
A recent, remarkable example of precision-driven initiative is provided by the DeltaCodes project~\cite{deltaTest2016},
where a grassroot community of code-owners struggled to extract well-defined computational results in the context of
Density Functional Theory (DFT) calculations. Ground-State quantum mechanical quantities as lattice constants and bulk modulus of
elementary crystals extracted from different codes were compared with each other, with the objective of reducing the uncertainties
on such data, and in some sense, to define a \emph{calibration} procedure for a Solid-State Physics DFT code.
Any other code able to extract the same quantities with the same approximations may be compared with the
results of the community in order to assess its computational reliability.
}

@@

In this paper, we focus our attention to Linear-Response treatments of \emph{molecular} systems,
in particular referring to LR-TDDFT calculation. We will investigate the impact that
\emph{open} (i.e. Isolated) Boundary Conditions (BC) that have to be imposed to the system will have on the computational evaluation of LR quantities.
\deleted[id=gen]{In order to do this}We perform analytic considerations on the equations beyond LR treatment and discuss the interplay between reproducible LR quantities and the \emph{locality} of the related objects. Such considerations will enable us to distinguish quantities that \textit{a priori} can be compared among different treatments - and therefore tested for reproducibility - from others that would \emph{explicitly} depend of the employed computational treatment.

\added[id=dal]{This analysis evidences that the description of molecular systems beyond the ground state unavoidably leads to the inclusion of delocalized
degrees of freedom above a specific energetic threshold. Moreover on the basis of this analysis we can assess something about the reliability of the results provided by computational setups able to explicit include the presence of delocalized degrees of freedom compared
to the ones in which these degrees of freedom are excluded by design.}

@@

\replaced[id=dal]{Let us start our discussion by reviewing some well-known concepts of quantum mechanic under a perspective that
will turn out to be very useful in the forthcoming considerations about linear-response quantities. In particular we revisit the impact
that the Isolated BC have on the solutions of the Schr\"odinger Equation.}{Let us start our discussion by revisiting the
impact that the Isolated BC have on the solutions of the time-independent Schr\"odinger Equation. We here present well-known concepts of
quantum mechanics, under a perspective that will turn out to be very useful in the forthcoming considerations about linear-response quantities.}

@@

On the other hand, positive energy solutions behave differently.
Even though the potential is localized \replaced[id=dal]{the associated Helmholtz kernel is a spherical wave and its}{, their positive energy are spherical waves and the} convolution with the source term provides a delocalized wavefunction $\psi(\r)$ in the whole space.
As they cannot be normalized they belong to the essential spectrum of the Hamiltonian operator and form a \emph{continuum} of states.
Outside of the support of the $\ket{\op V \psi}$ the solutions of Eq.~\eqref{SEHelmholtz1} can be put in bijection with the eigenstates of the
free-particle Hamiltonian $\op H_s$. For positive energies, the quantum numbers of the eigenstates are determined by the Laplacian operator,
regardless of the particular features of the potential.

@@

We can therefore introduce an explicit representation of the
response density parametrized as
\be\lb{rhoPrimeFluctuationStateDef1}
\dm'_\Phi(\omega) = \sum_p\left(\ketbra{\psi_p}{f_p^\Phi(-\omega)} + \ketbra{f_p^\Phi(\omega)}{\psi_p}\right) \;.
\ee
Here we have introduced the set of $\omega$-dependent \emph{fluctuation states} (FS)
$\ket{f_p^\Phi(\omega)} = \dm'_\Phi(\omega) \ket{\psi_p}$, that have values in the unoccupied subspace of $\hnot$.
\added[id=gen]{
However, non-hermiticity of $\dm'_\Phi(\omega)$ implies that the
fluctuation states are complex quantities.}
%  Causality implies that the real and imaginary part of FS are related each other by the Kramers-Kronig relations}
% \be\lb{KKrel}
% \mathrm{Re}\left( \ket{f_p^\Phi(\omega)} \right) =
% \mathcal P \int \dd \omega'
% \frac{\mathrm{Im}\left(\ket{f_p^\Phi(\omega')}\right)}{\omega -\omega'}.
% \ee
% \added[id=gen]{We will come to these equations in the forthcoming sections.}
%}

@@

\replaced[id=dal]{Equation \eqref{rhoPrimeFluctuationStateDef2} is the convenient starting point to assess the localization properties of the FS as it is easy to see that, for each choice of $p$, the value $\omega = |\eps_p|$ represents the threshold level that governs the asymptotic behavior of the associated Helmholtz kernel.}{It can be seen that for each choice of $p$, the value $\omega = |\eps_p|$ represents the threshold level
that governs the localization properties of the FS.}
As a ulterior \emph{stringent} hypothesis, let us now assume that the source term $\ket{s_p^\Phi(\omega)}$ is \added[id=dal]{spatially} localized, \textit{i.e.} its evaluation can be restricted on a bounded real-space domain.
Even when such locality argument holds true, for $\omega > |\eps_p|$,
the Helmholtz kernel behaves as a spherical wave and  gives raise to a $\ket{f^\Phi_p(\omega)}$ which is \emph{delocalized} in the entire real-space domain.
Such delocalization is an intrinsic consequence of the value of $\omega$, and does not depend on the particular choice of
$\op \Phi$.
Only for $\omega < |\eps_p|$ the kernel contains an exponential damping factor and $\ket{f^\Phi_p(\omega)}$ exhibits a bound-state behavior in the long range,
once again provided that the above locality arguments on $\ket{s_p^\Phi(\omega)}$ -- hence on $\op V'$ and $\Phi$ -- are valid.
\added[id=gen]{These arguments demonstate that for frequencies above the IP threshold energy, the
response density operator cannot be expressed in terms of localized states. Locality of
FS (and therefore of $\dm'_\Phi(\omega)$) can be only recovered for frequencies below such threshold.}

\deleted[id=dal]{Added a subsection for static perturbations}

@@

\deleted[id=dal]{Modified the name of the subsection}

\subsection{Localization propeties at finite $\omega$. Emergence of a strong and a weak regime}

@@

The analysis described above has shown that the building blocks of the response density $\{\ket{f_p(\omega)}\}$
are susceptible to exhibit bound-state long-range behavior \emph{only} when $\omega < |\eps_p|$,
while they behave as unbound states for values of $\omega$ above the threshold levels.
This fact has interesting implications in the evaluation of LR quantities. Indeed, writing the response
density in terms of FS implies that \deleted[id=dal]{(equation for the response density modified)}
% \be\lb{LinearResponseFunctDef2}
% \braket{\delta\op O}_\Phi(\omega) = 2 \mathrm{Re} \left(\sum_p \bra{\psi_p}\op O\ket{f_p^\Phi(\omega)}\right) \;.
% \ee
\begin{multline}\lb{LinearResponseFunctDef2}
\braket{\delta\op O}(\omega) = \sum_p \left( \bra{f_p^\Phi(-\omega)}\op O\ket{\psi_p}+ \bra{\psi_p}\op O\ket{f_p^\Phi(\omega)}\right) \\
=2 \sum_p\bra{\psi_p}\op O\ket{f_p^\Phi(\omega)}  \;.
\end{multline}
In the second equality we have assumed that $\op O$ is a real symmetric operator and chosen a real set of occupied molecular states.
Let us suppose that also the observable of interest $\op O$ \replaced[id=gen]{may be restricted to localized domains in the same way as the potentials $ \op V$ and $\op \Phi$.}{exhibits the locality properties introduced for the potentials $ \op V$, $\op \Phi$ and $\op V'$.}
Once again, it would be enough to consider a local operator for this condition to be met.

@@

On the contrary, the computation of the linear response functional can be much more demanding in the \emph{above threshold regime}, realized for $\omega>|\eps_h|$, in which fluctuation
states start behaving as unbound wave functions.
As by hypothesis $\ket{\op O \psi_p}$ is bound-state like, \emph{only} the scalar products of Eq.~\eqref{LinearResponseFunctDef2} can be evaluated in a localized domain.
In this case a computational description of fluctuation states is only meaningful in ``weak'' sense,
where the bound state character of $\{\bra{\psi_p}\op O\}$ behaves as a regulator.
This is naturally achieved in the above mentioned Krylov space treatments, which
explicitly deal with the expression of the scalar products.
\added[id=gen]{However, in this regime, imposing \emph{by design} a localized
behaviour of the FS might reveal very dangerous in view of convergence of the results:
the computational basis set have to be able to express in the support domain of $\{\bra{\psi_p}\op O\}$ a asymptotically delocalized FS. The degrees of completeness
of the computational basis should therefore be increased by imposing oscillatory degrees of freedom needed to represent the FS at the border of the simulation domain: it is important that the underlying basis set is able to express delocalised states.}

@@

As an example in support of these arguments we compare the \replaced[id=dal]{dynamical polarizability $\alpha(\omega)$}{absorption spectrum} of a $CO$ molecule computed in three distinct real-space computational setups, which differs for the dimension of the computational domain.
Results reported in \replaced[id=dal]{the two panels of figure \ref{co_spectrum} show the real and imaginary part of $\alpha(\omega)$, respectively. In all the cases a typical behavior emerges}{evidences a typical behavior}: the curves are almost coincident in the below threshold regime,
whereas the effect of the computational setup are clearly visible for values of  $\omega$ that exceed the threshold value.
For these high-frequency regions, convergence might be reached for domains of much larger sizes than what presented here~\cite{baroni2008},
which are -- as already noticed -- more than sufficient below threshold.

@@

% \begin{figure}
% \includegraphics[scale=0.6]{Fig2_CO_spectrum.eps}
% \caption{\label{co_spectrum}(Color online) Illustrative LR absorption spectra of the $CO$ molecule in computational setups with different sizes of the simulation domain.
% The value of the IP threshold is indicated by a vertical dashed line.}
% \end{figure}

@@

\added[id=dal]{It is possible to reformulate this type of analysis by directly looking at LR quantities that are defined independently from the perturbing operators and target observables. This can be achieved by studying the
\emph{susceptibility functional}, defined implicitly trough the operator equation}
\deleted[id=dal]{
In view of reproducibility, the scalar product of Eq.~\eqref{LinearResponseFunctDef1}, defining the observable fluctuation, may be compared among various numerical treatments.
However, even though directly related to experimental results, such quantity is not by itself a property of the molecule, as it is a function of $\omega$ that explicitly depends
on the perturbation operator $\op \Phi$, not to mention the target observable $\op O$. Alternatively, we may consider only the fluctuation state, but it is still a $\op \Phi$-dependent wavefunction,
with a potential local behavior only for $\omega$ below the ionization threshold. Hence, comparing fluctuation states among diverse numerical treatments would be a
challenging task.}
\deleted[id=dal]{It is therefore much more interesting to concentrate on LR quantities that are defined independently from the perturbing operators and target observables.
This can be achieved by studying the
\emph{susceptibility functional}, defined implicitly trough the operator equation}
\be\lb{SusceptibilityDef1}
\dm'(\omega) = \left(\omega -\mathcal L \right)^{-1} \commutator{\op \Phi(\omega)}{\dmnot} = \sop{{\cal \chi}}(\omega) \op\Phi(\omega) \;.
\ee
The susceptibility (super)operator $\cal \chi(\omega)$ is therefore closely related to the resolvent of the Liouvillian of Eq.~\eqref{LiouvillianRhopomegaDef1}, which is completely defined
from the molecule Hamiltonian $\hnot[\dm]$.
The ``excitation modes''  of the molecule are defined through the \emph{excitation operators}, satisfying
\be\lb{ExcitationOperatorsDef1}
\Liouv \op E_a = \Omega_a \op E_a \qq \op{\tilde E}_a \Liouv = \Omega_a \op{\tilde E}_a \;,
\ee
together with the operator orthonormalization condition %$\trace{\op{\tilde E}\op E'} = \delta(E-E')$.
\be\lb{orthoExcitatioOpDef1}
\trace{\op{\tilde E}_a\op E_b} = \delta_{ab} \;.
\ee
Excitations defined in that way may be considered as a basis of the operators that satisfy the transverse condition as in
Eq.~\eqref{RhopTransverseDef1}, and enable us to express the susceptibility as a spectral decomposition:
\be\lb{RhopExcitationDef1}
\sop{{\cal \chi}}(\omega) \cdot   = \sum_{\{a\}} \op E_{a} \,
\frac{\trace{\commutator{\op{\tilde E}_a}{\dmnot}\cdot}}{\omega-\Omega_a} \;.
\ee
In this way Eq.~\eqref{LinearResponseFunctDef2} may be also written as:
\begin{equation}\lb{deltaoexc}
  \braket{\delta \op O} =  \sum_{\{a\}} \frac{\trace{\op O \sop B_a \op \Phi(\omega)}}{\omega -\Omega_a}
\end{equation}
where the action of the spectral operator $\sop B_a$ reads
\begin{equation}
  \trace{\op O \sop B_a \op \Phi(\omega)} = \trace{\op O \op E_a}
  \trace{\op{\tilde E}_a \commutator{\dmnot}{\op \Phi(\omega)}}
\end{equation}

>>>>

% Useful packages
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage{mathtools,braket}
\usepackage{tikz}
\usepackage{xifthen}
%\pgfplotsset{compat=1.15}
\usepackage[markup=underline]{changes}
%\usepackage[final]{changes}
\definechangesauthor{dal}
\definechangesauthor{gen}

@@

\section{Introduction}

% Since the days of their foundations, disciplines like Computational Physics or Quantum Chemistry always had to deal with the problem of the
% computational reliability of the results. More specifically, a computational result may be considered reliable for various reasons.
% For example, it may be in agreement with some previous observation or measurement that the employed computational model is susceptible to reproduce.
% The model itself may in this way be classified as ``accurate''; it can then be employed to extract, \textit{in silico},
% other quantities that might be then used to \emph{predict} or \emph{interpret} experiments.
% Another concept of reliability is the concept of  ``precision'', which is more focused to the actual results
% that a given computational model provides for a specific set of quantities, regardless of their potential agreement with
% experimental data. The aim of precise computational approach is to reduce the \emph{computational} uncertainties
% of quantities extracted out of a well-defined model, and provide reference results, to which other computer codes
% employing the same model can be compared.
% 
% A recent, remarkable example of precision-driven initiative is provided by the DeltaCodes project~\cite{deltaTest2016},
% where a grassroot community of code-owners struggled to extract well-defined computational results in the context of
% Density Functional Theory (DFT) calculations. Ground-State quantum mechanical quantities as lattice constants and bulk modulus of
% elementary crystals extracted from different codes were compared with each other, with the objective of reducing the uncertainties
% on such data, and in some sense, to define a \emph{calibration} procedure for a Solid-State Physics DFT code.
% Any other code able to extract the same quantities with the same approximations may be compared with the
% results of the community in order to assess its computational reliability.

@@

In this paper, we focus our attention to Linear-Response treatments of \emph{molecular} systems,
in particular referring to LR-TDDFT calculation. We will investigate the impact that
\emph{open} (i.e. Isolated) Boundary Conditions (BC) that have to be imposed to the system will have on the computational evaluation of LR quantities.
We perform analytic considerations on the equations beyond LR treatment and discuss the interplay between reproducible LR quantities and the \emph{locality} of the related objects. Such considerations will enable us to distinguish quantities that \textit{a priori} can be compared among different treatments - and therefore tested for reproducibility - from others that would \emph{explicitly} depend of the employed computational treatment.

This analysis evidences that the description of molecular systems beyond the ground state unavoidably leads to the inclusion of delocalized degrees of freedom above a specific energetic threshold. Moreover on the basis of this analysis we can assess something about the reliability of the results provided by computational setups able to explicit include the presence of delocalized degrees of freedom compared
to the ones in which these degrees of freedom are excluded by design.

@@

% {Let us start our discussion by revisiting the
% impact that the Isolated BC have on the solutions of the time-independent Schr\"odinger Equation. We here present well-known concepts of
% quantum mechanics, under a perspective that will turn out to be very useful in the forthcoming considerations about linear-response quantities.}
Let us start our discussion by reviewing some well-known concepts of quantum mechanic under a perspective that
will turn out to be very useful in the forthcoming considerations about linear-response quantities. In particular we revisit the impact that the Isolated BC have on the solutions of the Schr\"odinger Equation.

@@

On the other hand, positive energy solutions behave differently.
Even though the potential is localized the associated Helmholtz kernel is a spherical wave and its convolution with the source term provides a delocalized wavefunction $\psi(\r)$ in the whole space.
As they cannot be normalized they belong to the essential spectrum of the Hamiltonian operator and form a \emph{continuum} of states.
Outside of the support of the $\ket{\op V \psi}$ the solutions of Eq.~\eqref{SEHelmholtz1} can be put in bijection with the eigenstates of the
free-particle Hamiltonian $\op H_s$. For positive energies, the quantum numbers of the eigenstates are determined by the Laplacian operator,
regardless of the particular features of the potential.

@@

We can therefore introduce an explicit representation of the
response density parametrized as
\be\lb{rhoPrimeFluctuationStateDef1}
\dm'_\Phi(\omega) = \sum_p\left(\ketbra{\psi_p}{f_p^\Phi(-\omega)} + \ketbra{f_p^\Phi(\omega)}{\psi_p}\right) \;.
\ee
Here we have introduced the set of $\omega$-dependent \emph{fluctuation states} (FS)
$\ket{f_p^\Phi(\omega)} = \dm'_\Phi(\omega) \ket{\psi_p}$, that have values in the unoccupied subspace of $\hnot$.
However, non-hermiticity of $\dm'_\Phi(\omega)$ implies that the fluctuation states are complex quantities.
%  Causality implies that the real and imaginary part of FS are related each other by the Kramers-Kronig relations}
% \be\lb{KKrel}
% \mathrm{Re}\left( \ket{f_p^\Phi(\omega)} \right) =
% \mathcal P \int \dd \omega'
% \frac{\mathrm{Im}\left(\ket{f_p^\Phi(\omega')}\right)}{\omega -\omega'}.
% \ee
% \added[id=gen]{We will come to these equations in the forthcoming sections.}
%}

@@

Equation \eqref{rhoPrimeFluctuationStateDef2} is the convenient starting point to assess the localization properties of the FS as it is easy to see that, for each choice of $p$, the value $\omega = |\eps_p|$ represents the threshold level that governs the asymptotic behavior of the associated Helmholtz kernel.
As a ulterior \emph{stringent} hypothesis, let us now assume that the source term $\ket{s_p^\Phi(\omega)}$ is spatially localized, \textit{i.e.} its evaluation can be restricted on a bounded real-space domain.
Even when such locality argument holds true, for $\omega > |\eps_p|$,
the Helmholtz kernel behaves as a spherical wave and  gives raise to a $\ket{f^\Phi_p(\omega)}$ which is \emph{delocalized} in the entire real-space domain.
Such delocalization is an intrinsic consequence of the value of $\omega$, and does not depend on the particular choice of
$\op \Phi$.
Only for $\omega < |\eps_p|$ the kernel contains an exponential damping factor and $\ket{f^\Phi_p(\omega)}$ exhibits a bound-state behavior in the long range,
once again provided that the above locality arguments on $\ket{s_p^\Phi(\omega)}$ -- hence on $\op V'$ and $\Phi$ -- are valid.
These arguments demonstrate that for frequencies above the IP threshold energy, the
response density operator cannot be expressed in terms of localized states. Locality of
FS (and therefore of $\dm'_\Phi(\omega)$) can be only recovered for frequencies below such threshold.

@@

\subsection{Localization properties at finite $\omega$. Emergence of a strong and a weak regime}

@@

The analysis described above has shown that the building blocks of the response density $\{\ket{f_p(\omega)}\}$
are susceptible to exhibit bound-state long-range behavior \emph{only} when $\omega < |\eps_p|$,
while they behave as unbound states for values of $\omega$ above the threshold levels.
This fact has interesting implications in the evaluation of LR quantities. Indeed, writing the response
density in terms of FS implies that
% \begin{align}\lb{LinearResponseFunctDef2}
% \braket{\delta\op O}_\Phi(\omega) = &\sum_p \left( \bra{f_p^\Phi(-\omega)}\op O\ket{\psi_p}+ \bra{\psi_p}\op O\ket{f_p^\Phi(\omega)}\right) \nn \\
% = &2 \sum_p\bra{\psi_p}\op O\ket{f_p^\Phi(\omega)}  \;,
% \end{align}
\be\lb{LinearResponseFunctDef2}
\braket{\delta\op O}_\Phi(\omega) = 2 \sum_p \bra{\psi_p}\op O\ket{f_p^\Phi(\omega)} \;,
\ee
where we have assumed that $\op O$ is a real symmetric operator and chosen a real set of occupied molecular states.
Let us suppose that also the observable of interest $\op O$ may be restricted to localized domains in the same way as the potentials $ \op V$ and $\op \Phi$. Once again, it would be enough to consider a local operator for this condition to be met.

@@

On the contrary, the computation of the linear response functional can be much more demanding in the \emph{above threshold regime}, realized for $\omega>|\eps_h|$, in which fluctuation
states start behaving as unbound wave functions.
As by hypothesis $\ket{\op O \psi_p}$ is bound-state like, \emph{only} the scalar products of Eq.~\eqref{LinearResponseFunctDef2} can be evaluated in a localized domain.
In this case a computational description of fluctuation states is only meaningful in ``weak'' sense,
where the bound state character of $\{\bra{\psi_p}\op O\}$ behaves as a regulator.
This is naturally achieved in the above mentioned Krylov space treatments, which
explicitly deal with the expression of the scalar products.
However, in this regime, imposing \emph{by design} a localized
behaviour of the FS might reveal very dangerous in view of convergence of the results:
the computational basis set have to be able to express in the support domain of $\{\bra{\psi_p}\op O\}$ a asymptotically delocalized FS. The degrees of completeness
of the computational basis should therefore be increased by imposing oscillatory degrees of freedom needed to represent the FS at the border of the simulation domain: it is important that the underlying basis set is able to express delocalised states

@@

As an example in support of these arguments we compare the dynamical polarizability $\alpha(\omega)$ of a $CO$ molecule computed in three distinct real-space computational setups, which differs for the dimension of the computational domain.
Results reported in the two panels of figure \ref{co_spectrum} show the real and imaginary part of $\alpha(\omega)$, respectively. In all the cases a typical behavior emerges: the curves are almost coincident in the below threshold regime,
whereas the effect of the computational setup are clearly visible for values of  $\omega$ that exceed the threshold value.
For these high-frequency regions, convergence might be reached for domains of much larger sizes than what presented here~\cite{baroni2008}, which are -- as already noticed -- more than sufficient below 
threshold\footnote{Add the convergence domain for the benzene molecule by looking at the results of...}.

@@

@@

It is possible to reformulate this type of analysis by directly looking at LR quantities that are defined independently from the perturbing operators and target observables. This can be achieved by studying the
\emph{susceptibility functional}, defined implicitly trough the operator equation
% In view of reproducibility, the scalar product of Eq.~\eqref{LinearResponseFunctDef1}, defining the observable fluctuation, may be compared among various numerical treatments.
% However, even though directly related to experimental results, such quantity is not by itself a property of the molecule, as it is a function of $\omega$ that explicitly depends
% on the perturbation operator $\op \Phi$, not to mention the target observable $\op O$. Alternatively, we may consider only the fluctuation state, but it is still a $\op \Phi$-dependent wavefunction,
% with a potential local behavior only for $\omega$ below the ionization threshold. Hence, comparing fluctuation states among diverse numerical treatments would be a
% challenging task.
% It is therefore much more interesting to concentrate on LR quantities that are defined independently from 
% the perturbing operators and target observables. This can be achieved by studying the \emph{susceptibility
%functional}, defined implicitly trough the operator equation
\be\lb{SusceptibilityDef1}
\dm'(\omega) = \left(\omega -\mathcal L \right)^{-1} \commutator{\op \Phi(\omega)}{\dmnot} = \sop{{\cal \chi}}(\omega) \op\Phi(\omega) \;.
\ee
The susceptibility (super)operator $\cal \chi(\omega)$ is therefore closely related to the resolvent of the Liouvillian of Eq.~\eqref{LiouvillianRhopomegaDef1}, which is completely defined
from the molecule Hamiltonian $\hnot[\dm]$.
The ``excitation modes''  of the molecule are defined through the \emph{excitation operators}, satisfying
\be\lb{ExcitationOperatorsDef1}
\Liouv \op E_a = \Omega_a \op E_a \qq \op{\tilde E}_a \Liouv = \Omega_a \op{\tilde E}_a \;,
\ee
together with the operator orthonormalization condition %$\trace{\op{\tilde E}\op E'} = \delta(E-E')$.
\be\lb{orthoExcitatioOpDef1}
\trace{\op{\tilde E}_a\op E_b} = \delta_{ab} \;.
\ee
Excitations defined in that way represent an operator basis that satisfy the transverse condition as in
Eq.~\eqref{RhopTransverseDef1} (see appendix \ref{LiouvillianAction}), and enable us to express the susceptibility as a spectral decomposition:
\be\lb{RhopExcitationDef1}
\sop{{\cal \chi}}(\omega) \cdot   = \sum_{\{a\}} \op E_{a} \,
\frac{\trace{\commutator{\op{\tilde E}_a}{\dmnot}\cdot}}{\omega-\Omega_a} \;.
\ee
In this way Eq.~\eqref{LinearResponseFunctDef2} may be also written as:
\begin{equation}\lb{deltaoexc}
  \braket{\delta \op O} =  \sum_{\{a\}} \frac{\trace{\op O \sop B_a \op \Phi(\omega)}}{\omega -\Omega_a}
\end{equation}
where the action of the spectral operator $\sop B_a$ reads
\begin{equation}
  \trace{\op O \sop B_a \op \Phi(\omega)} = \trace{\op O \op E_a}
  \trace{\op{\tilde E}_a \commutator{\dmnot}{\op \Phi(\omega)}}
\end{equation}
